---
title: "Netflix Engagement Report: A Case Study"
author: "May Eindra Tet Toe"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.align = "center", out.width = "85%")
```

---

# Summary

Insert summary here

**Key findings:**

---

# Introduction

Netflix publishes a bi-annual engagement report disclosing viewership data 
for its titles. This transparency gives analysts a rare look into what audiences 
actually watch — not just what is trending. The dataset is sourced from 
TidyTuesday (July 2025), and it includes view counts and hours viewed for movies and shows, 
along with release dates, runtimes, and global availability flags.

This analysis follows the six phases of the data analysis process:
 **Ask, Prepare, Process, Analyze, Share, and Act.**

---

# Ask: Business Task

**Central question:** What patterns in Netflix's 2025 engagement data can help explain content performance, and what do those patterns suggest about audience behavior?

Supporting questions:

1. Which titles drove the most engagement by hours viewed?
2. How do Movies and TV Shows compare in total viewership?
3. Does global availability significantly impact view counts and hours?
4. Is there a relationship between content runtime and engagement?
5. What release timing patterns are associated with higher viewership?

---

# Prepare: Data Source

The data comes from Netflix's official engagement report, compiled by the
 [TidyTuesday project](https://github.com/rfordatascience/tidytuesday/tree/main/data/2025/2025-07-29).

| File | Description |
|------|-------------|
| `movies.csv` | Netflix movie titles with viewership metrics |
| `shows.csv` | Netflix TV show titles with viewership metrics |

**Columns (both files):**

| Column | Description |
|--------|-------------|
| `source` | Source file name |
| `report` | Report period |
| `title` | Movie/Show title |
| `available_globally` | Whether the show/movie is available globally or not |
| `release_date` | Date the show/movie was released on Netflix |
| `hours_viewed` | Total hours viewed on the show/movie |
| `runtime` | Content runtime |
| `views` | Total number of hours viewed |

**Limitations:** The report covers only Jan–Jun 2025 and excludes titles that did not 
reach a minimum viewership threshold. It does not include genre, language, or production 
budget data.

---

# Process: Data Cleaning & Preparation

## Load Libraries

```{r load-libraries}
library(tidyverse)
library(knitr)
library(scales)
```

## Load Data

```{r load-data}
movies <- readr::read_csv(
  'https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-07-29/movies.csv')

shows <- readr::read_csv(
  'https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-07-29/shows.csv')
```

## Initial Inspection

```{r inspect-data}
glimpse(movies)
glimpse(shows)
```

```{r check-missing}
colSums(is.na(movies))
colSums(is.na(shows))
```

## Parse Runtime to Minutes

The `runtime` column is stored as a string (e.g., `"1H 54M"`). We convert it to numeric minutes for analysis.

```{r parse-runtime}
parse_runtime_minutes <- function(runtime_str) {
    hours   <- as.numeric(str_extract(runtime_str, "(?<=^|\\s)(\\d+)(?=H)"))
    minutes <- as.numeric(str_extract(runtime_str, "(\\d+)(?=M)"))
    hours   <- replace_na(hours, 0)
    minutes <- replace_na(minutes, 0)
    hours * 60 + minutes
    }
```

## Combine Datasets


```{r combine-data}
movies_clean <- movies |>
mutate(
    content_type    = "Movie",
    runtime_minutes = parse_runtime_minutes(runtime),
    release_date    = as.Date(release_date),
    release_month   = month(release_date, label = TRUE, abbr = TRUE),
    release_quarter = paste0("Q", quarter(release_date)),
    global          = if_else(available_globally == "Yes", "Global", "Non-Global")
    )


shows_clean <- shows |>
mutate(
    content_type    = "TV Show",
    runtime_minutes = parse_runtime_minutes(runtime),
    release_date    = as.Date(release_date),
    release_month   = month(release_date, label = TRUE, abbr = TRUE),
    release_quarter = paste0("Q", quarter(release_date)),
    global          = if_else(available_globally == "Yes", "Global", "Non-Global")
    )

all_content <- bind_rows(movies_clean, shows_clean)
```

